### 2022-06-27 paper/Chapters/Chapter01.tex
update fugrc, 5 times downsampling

Surface normal is an important property of a surface with many applications, like surface reconstruction, shadings generation and other visual effects. The quality of the surface normal is directly impact the performances of the techniques mentioned above. However, especially in the task of real-world object digitalization, the surface is usually hardly to be mathematically described in equations due to the elaborate details on the objects. Instead, it is common to use a group of points to describe the object surface, which is a memory economical solution and also can be easier measured by 3D scanners.  The working principle of scanners are various due to the application scenarios, which consequently produce different point cloud structure. For the scanners without positions recording, the point cloud acquired after scanning is unstructured. In this case, every 3D point can be captured by different capture position, and neighbors are not defined by capture time, which increase the difficulty and computation for the neighbor based normal inference approaches.  Furthermore, since the lack of inherent structure, the normal can hardly be inferred by the parallel approaches. 
To acquire the structured point cloud, a depth camera can be used for data collection. It captures the RGB-D images for the object, which includes the standard RGB image with depth information of each pixel. After camera calibration, the corresponding point cloud can be calculated based on the depth map and camera matrix. The advantages are, a structured point cloud is mapped directly based on the same capture position from a 2D depth map, the neighbor information of each point is identical to the corresponding pixel in the 2D depth map, which reduced a huge computational works. Besides, base on the RGB image, which has the same indices with depth map, the normal inference task can also utilize the relationship between surface normal and light reflection to further rectify the inferred normal map with knowing light source. Furthermore, the structured point cloud has the same shape with the standard RGB images, which can be used for deep learning based approaches. 
Unfortunately, the depth maps captured by the sensors are only semi-dense, which is mainly caused by optical noises and the reflections in dark and shinny areas. Consequently, it raises the challenges to the accuracy of normal inference tasks. 
Recently, deep learning-based methods are used in computer vision tasks such as image segmentation, image inpainting, and depth density enhancement. The output of the model is usually has the same shape as the input but with different content, which is by comparison similar to the normal inference task. 
The deep learning based method has many advantages, first, it deals well with semi-dense input, which is a common issue for the depth map. Second, the network of deep learning model handles the structured point cloud as a single input and infer the corresponding normal all together, which is very time economical comparing to the approaches like neighbor based methods. Furthermore, if the texture image acquiring from the camera is already illuminated by strong directional light of a video projector, whose position is known, then there should exist theoretical relations among light direction, normal direction, and grayscale image. Then the normal can be inferred better using the given image information and depth map.  Based on the grayscale image and corresponding semi-dense depth maps, a multi-staged model can be designed to infer the normal map, which can give more density and robust results comparing to the standard algorithm.
In this work, we apply the deep learning based approach for surface normal inference, which is based on the calibrated illuminated RGB-D image. A network named gated convolution neural network (GCNN) is proposed for normal inference, which is robust to the semi-dense input. 
With the help of synthetic data in Unity, a dataset is also created for CNN model training. It can provide accurate ground truth for training work, which real data is usually not provided. The trained model is performed well both in synthetic dataset as well as the real dataset captured from RGB-D cameras and achieve a remarkably better prediction accuracy at a low computational cost compared to the standard approaches for semi-dense point clouds.  

>>>>

Surface normal is an important property of a surface with many applications, like surface reconstruction, shadings generation and other visual effects. However, the calculation of surface normal in many tasks is not as straightforward as simply the cross-product of two plane vectors. Especially in the task of real-world object digitalization, the surface is usually hardly to be mathematically described in equations due to the elaborate details on the objects. Instead, it is common to use a group of points to describe the object surface, which is a memory economical solution to save the fine detail of the objects and also can be easier measured by 3D scanners. Therefore the task is converted to the normal inference based on the surface point, and the most of the case, the result surface normal is merely an approximation.

\begin{figure}[h!]
	\centering
	{\includegraphics[width=.45\textwidth]{./Figures/point-cloud.png}};
	{\includegraphics[width=.45\textwidth]{./Figures/point-cloud-zoom-in.png}};
	\caption{Left: A part of the point cloud of the dragon model. Right: The zoom in of the left point cloud.}
	\label{fig:point-cloud}
\end{figure}

Apart from this, due to the application scenarios, the working principle of scanners are various, which consequently produce point cloud structures in different forms. For the scanners without positions recording, the point cloud acquired after scanning is unstructured. In this case, every 3D point can be captured by different capture position, and neighbors are not defined by capture time. It increases the difficulty and computation for the neighbor based normal inference approaches.  Furthermore, since the lack of inherent structure, the normal can hardly be inferred by the parallel approaches.

However, for the scanner with calibrated camera, a structured point cloud can be captured. One example is the depth camera. It captures the RGB-D images for the object, which includes the standard RGB image with depth information of each pixel. After camera calibration, a corresponding point cloud can be calculated based on the depth map and camera matrix. It gives the advantage that a structured point cloud is mapped directly based on the same capture position from a 2D depth map, the neighbor information of each point is identical to the corresponding pixel in the 2D depth map. It provides a better view for the normal inference task since the neighbor information can be considered as an important reference.
Besides the point cloud, which is one piece of information can be used for normal reference, the RGB image also leaves the trace of surface normal. If assume object surface is diffuse and the light coming from a knowing direction , the brightness of the object surface is proportional to the surface normal, which is known as Lambertian surface. It indicates the normal inference task can also utilize the relationship between surface normal and light reflection to further rectify the inferred normal map with knowing light source. Therefore, the illuminated calibrated RGB-D image provides as a good input for normal inference tasks.
Unfortunately, in the actual situation, the depth maps captured by the sensors are only semi-dense, which is mainly caused by optical noises and the reflections in dark and shinny areas. Consequently, it disrupts the robustness of the normal inference methods. Median filters can be used for the sparse missing pixels, however, for the case of huge missing holes in the depth map, it produces just a paltry result. Thus a reasonable guess is required for missing areas.

Deep learning based methods provides multiple possible solutions for the challenges mentioned above. First, to deal with the noised input, deep learning based methods already have the solution for the similar tasks like image inpainting and depth density enhancement, which base on the noised image as input to predict the clean and fully dense output with or without original meanings. Therefore, it can be a help for noise in the depth map. Besides, the network of deep learning model can also handle the structured point cloud as a single input and infer the corresponding normal all together, which is very time economical comparing to the approaches like neighbor based methods. In addition, the multi-stages training architectures provide a way to consider both depth map and texture image as the input to predict the surface normals, which not only consider the point cloud but also the add the lambertian reflection as a further constraint.

In this work, we focus on the normal inference based on a semi-dense depth map with RGB image using deep learning based approaches. Specifically, the missing pixels in the depth map is filled up by the gated convolution and propagate in a customized UNet with skipping connections. The output of the training model is directly the surface normal corresponding the input depth map. The grayscale image is used to further improve the estimation accuracy. 
For the training work, a dataset named “synthetic50-5” is created including 55 high resolution point clouds from internet, as shown in Appendix A. The point clouds provide with the high accurate normals, which can be used as the ground truth of the training work. Most of the models have elaborate details with high curvatures but also contain smooth surfaces. The trained model is evaluated on both synthetic dataset as well as the real dataset captured from RGB-D cameras. A series of metrics have been used for qualitative evaluation. The model is shown to achieve a remarkably better prediction accuracy at a low computational cost compared to the standard approaches for semi-dense point clouds.

### 2022-06-27 paper/Chapters/Chapter02.tex
update fugrc, 5 times downsampling

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
%-----------------------------------
%	SUBSECTION 1
%-----------------------------------
\section{Standard Methods}

%%%%%%%%%%%%%%%%%%% talk about the reason of this thesis %%%%%%%%%%%%%%%%%%

%% standard method for normal inference is insufficient, input is sparse, not robust enough
Standard methods compute normals from the point cloud using neighboring information in image space or from a single grayscale image using use Shape from Shading \cite{SFS}. The first method assumes that the neighbors of the points locate on the same plane. This method performs well with a well-chosen window size. However, the drawbacks are that the algorithm is highly noise sensitive. It is weak in handling missing pixels, which is a common issue in the input data. The second method depends on the correct information about the light source. Errors may occur in regions with inter-reflections in the 3D measurement.

@@

\section{Deep Learning based Method}
%% CNN based methods
Recently, deep learning based method \cite{yolov3}, \cite{efficientDet} achieved a great succeed for image processing. 
These network architectures use a batch of RGB/Grayscale images as input and usually employ for classification problems. Usually, the images are convoluted with a convolutional layer and downsampling with pooling layers. The outputs of the network consist of a single value to represent the index of the corresponding class \cite{efficientDet}, or with a set of values to represent the position of bounding boxes.\cite{yolov3}.

However, in many other vision tasks, like normal map inference, the output is demanded as the same shape as the input. Instead of predicting one or several classes for the whole input matrix, the class for each pixel requires for prediction. In this case, the traditional network architecture is not suitable anymore.

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------

@@

\section{Sparse Input processing}

%% methods deal with sparse input, like nconv, gconv ...
The depth map is supposed to be dense. Therefore, how to accept sparse depth map as input in CNNs is one of the most important problem. Some trivial solutions like median filters are good enough, if the missing pixels are sparse enough, however, for the case of huge missing holes in the depth map, it produces just a paltry result. Thus a reasonable guess is required for missing areas. Generally, it can be solved as image inpainting problems,\cite{inpainting1},\cite{inpainting2}.

%% Deep learning based approaches
Some deep learning based method for image inpainting also achieved quite good performance for the hole mending task. 
%% talk about gconv
Notably, in 2016, Oord et al. \cite{gated_activation} proposed a gated activation unit for a CNN model,
\[\textbf{y} = \tanh (W_{k,f} * \textbf{x}) \odot \sigma (W_{k,g} * \textbf{x})\]
to substitute the standard activation layer, where $ \sigma $ is the sigmoid function, which constricts the output value of second part between $ [0,1] $.  The function is inspired by Long Short-Term Memory (LSTM) \cite{lstm} and Rated Recurrent Unit (GRU).\cite{gru} It is originally used for learning complex interactions as LSTM gates does. In 2018, Yu et al. \cite{gconv} employed same function for free-form image inpainting, which can be used to learn mask automatically from image it self.

%% talk about nconv
Different to aforementioned approaches, Knutsson et al. in 2005 introduced normalized convolution \cite{nconv} dealing with missing sample case for convolution operation, which aims to reconstruct the missing pixels from the sparse output sensed by cameras, which particularly considered the confidence of each interpolated pixels, since it provides the trustworthiness of the predicted value. The higher the reliability of the value inference, the better the model shape reconstruction.

In 2018, Eldesokey et al. \cite{ncnn} applied normalized convolution in CNN as normalized convolution layer that takes both sparse depth map and a binary confidence map as input to perform scene depth completion.
In 2020, Eldesokey et al. \cite{pncnn} focus on modeling the uncertainty of depth data instead of assuming binary input confidence.

Guided method \cite{guided} requires addition information like RGB image or the certainty map of depth map, and fuse them together to predict the dense depth map.

@@

%% CNN based methods
Recently, CNN based methods improve the performance of image processing to a brand new stage. In 2014, Eigen et al.\cite{Eigen} proposed a method predicting depth map directly from RGB image using CNN. In this case, no depth map is required. In 2016, Laina et al. \cite{img2depth} proposed a deeper network based on ResNet \cite{resnet} with a well designed upsamling part. 
%% guided scene completion
In 2018, Qi et al. proposed GeoNet\cite{GeoNet}, it integrates both algebra method and also CNN method to inference depthmap based on \cite{img2depth} \cite{geometry_based_solution}.

%%\begin{figure}[!h]
%%	\centering
%%	\includegraphics[width=\textwidth]{./ref/u-net-illustration-correct-scale2}
%%	\caption{U-net architecture (example for 32x32 pixels in the lowest resolution). Each blue box corresponds to a multi-channel feature map. The number of channels is denoted on top of the box. The x-y-size is provided at the lower left edge of the box. White boxes represent copied feature maps. The arrows denote the different operations.
	%%	}
%%	\label{fig:u-net}
%% \end{figure}

%% unstructured point cloud 
In some case, the input is unstructured point cloud which can not be fed into a CNN entirely. Thus, a challenge task connect to the deep learning is the input format. Since different point clouds have different sizes. 
In 2018, Ben-Shabat et al. \cite{Ben-Shabat_2019_CVPR} presented Nesti-Net. It predicts the normal point by point with the help of neighbor points. It fixed the distance of considering neighbors to provide an unified input for CNN. In 2021, Zhou et al. \cite{zhou2021fast} presents a method considering overlapping of different patches (a group of neighboring points) as input to evaluate normals.

>>>>

Guided method \cite{guided} requires addition information like RGB image for the inference.

%% denoising
The depth map captured by depth sensors are usually with full of missing pixels and holes.
\cite{depth-inpainting-distribution} proposed a depth hole filling method using the depth distributions of neighboring regions.

\cite{nconv} introduced normalized convolution dealing with missing or uncertain data for convolution operation. It uses a binary mask to distinguish missing data and integrate it into the convolution operation. \cite{ncnn} applied it and use normalized convolution layers in their networks, which aims to reconstruct the missing pixels from the sparse depth map sensed by cameras. \cite{pncnn} proposed an input confidence estimation network to estimate the confidence instead of using a binary mask. However, the relevant papers are only using 1 channel data as model input, which didn't discussed the case for the multiple-channel data as input.

\cite{gated_activation} proposed a gated activation unit to model more complex interactions comparing to standard CNN layers, which mainly inspired by the multiplicative units exist in Long Short-Term Memory proposed by \cite{lstm} and Rated Recurrent Unit (GRU) proposed by \cite{gru}. 
\cite{gconv} employed the same gated unit solving for the free-form image inpainting task. The proposed network use 3 channel RGB images as input and estimate the missing pixels.

%% method for structured point cloud

%% CNN based methods

estimate the surface normal based on improve the performance of image processing to a brand new stage. 
In 2014, Eigen et al.\cite{Eigen} proposed a method predicting depth map directly from RGB image using CNN. In this case, no depth map is required. In 2016, Laina et al. \cite{img2depth} proposed a deeper network based on ResNet \cite{resnet} with a well designed upsamling part. 
%% guided scene completion

\cite{geometry_based_solution} proposed a method to learn discriminative and geometrically informative primitives from RGBD images, which is further used to recover the surface normals of a scene from a single image.

\cite{GeoNet} uses ResNet proposed by \cite{resnet} to infer the surface normal from RGB image, and further refine it with the help of depth map based on the methods proposed by \cite{geometry_based_solution}  further refine the a  GeoNet to infer depth and surface normal in one unified system,  method proposed by \cite{img2depth}.

%% unstructured point cloud
For the unstructured point cloud, the neighbor information of each point is usually unknown. K-nearest neighbor is a common algorithm for neighbor searching. With knowing this information, the neighbor based approaches can be used as a second step. However, KNN-method merely based on the Euclidean distance in the 3D space. Therefore, the points of the other surfaces but in a close distance will also be considered as neighbors. \cite{unstructed-pc} proposed a method based on unstructured point cloud with selecting the optimal scale around each point for normal estimation. \cite{unstructed-pc-patch-stitching} presents a pipeline for calculating the normals of multiple points in parallel.

\section{Standard Methods}

Standard methods compute normals from the point cloud using neighboring information in image space or from a single grayscale image using use Shape from Shading \cite{SFS}. The first method assumes that the neighbors of the points locate on the same plane. This method performs well with a well-chosen window size. However, the drawbacks are that the algorithm is highly noise sensitive. It is weak in handling missing pixels, which is a common issue in the input data. The second method depends on the correct information about the light source. Errors may occur in regions with inter-reflections in the 3D measurement.

@@

\section{Deep Learning based Method}
%% CNN based methods

Recently, deep learning based method achieved a great succeed for image processing.  \cite{yolov3} proposed  \cite{efficientDet} These network architectures use a batch of RGB/Grayscale images as input and are employed for classification problems. Usually, the images are convoluted with a convolution layer and downsampling with pooling layers. The outputs of the network consist of a single value to represent the index of the corresponding class \cite{efficientDet}, or with a set of values to represent the position of bounding boxes.\cite{yolov3}. However, in many other vision tasks, like normal map inference, the output is demanded as the same shape as the input. Instead of predicting one or several classes for the whole input matrix, the class for each pixel requires for prediction. In this case, the traditional network architecture is not suitable anymore.

@@

@@
